Model definition

### 2025-05-15 nbs/tutorials/models.qmd
docs(tutorials): rejoin models code blocks and print model structure
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

```python
import torch
import pyro
import anndata as ad
import numpy as np
import matplotlib.pyplot as plt
from pyrovelocity.utils import print_anndata, anndata_string, print_string_diff
from pyrovelocity.models.modular import PyroVelocityModel
from pyrovelocity.models.modular.factory import create_legacy_model1

# Set random seed for reproducibility
torch.manual_seed(42)
np.random.seed(42)
pyro.set_rng_seed(42)
```

```python
# Create synthetic data
n_cells, n_genes = 100, 10
u_data = np.random.poisson(5, size=(n_cells, n_genes))
s_data = np.random.poisson(5, size=(n_cells, n_genes))

@@

# Set up AnnData for PyroVelocity
adata = PyroVelocityModel.setup_anndata(adata)
final_data_state_representation = anndata_string(adata)
print_string_diff(
    initial_data_state_representation,
    final_data_state_representation,
    diff_title="Setup AnnData diff",
)
print_anndata(adata)
```

```python
# Create a standard model with default components
model = create_legacy_model1()
```

```python
# Train the model
model.train(
    adata=adata,
    max_epochs=20,  # Use a small number for demonstration
    learning_rate=0.01,
    use_gpu=False  # Set to True if GPU is available
)
```

```python
# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=30  # Use a small number for demonstration
)
```

```python
# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)

# Compute velocity and store it in AnnData
velocity = model.get_velocity(
    adata=adata,
    num_samples=30,
    basis='umap'  # This will be used for embedding
)
```

```python
# Print available keys in posterior samples
print("Available keys in posterior samples:")
for key in posterior_samples.keys():
    print(f"- {key}: {posterior_samples[key].shape}")
```

@@

# Visualize velocity
scv.pl.velocity_embedding_stream(
    adata,
    basis='umap',
    color='gene_0',
    vkey='velocity_pyro'  # Use the velocity key added by get_velocity
)
```

>>>>

```python
import torch
import pyro
import anndata as ad
import numpy as np
import matplotlib.pyplot as plt
import scanpy as sc
from pyrovelocity.utils import print_anndata, anndata_string, print_string_diff
from pyrovelocity.models.modular import PyroVelocityModel
from pyrovelocity.models.modular.factory import create_legacy_model1

# Set random seed for reproducibility
torch.manual_seed(42)
np.random.seed(42)
pyro.set_rng_seed(42)

# Create synthetic data
n_cells, n_genes = 100, 10
u_data = np.random.poisson(5, size=(n_cells, n_genes))
s_data = np.random.poisson(5, size=(n_cells, n_genes))

@@

# Set up AnnData for PyroVelocity
adata = PyroVelocityModel.setup_anndata(adata)
post_setup_anndata_representation = anndata_string(adata)
print_string_diff(
    initial_data_state_representation,
    post_setup_anndata_representation,
    diff_title="Setup AnnData diff",
)
print_anndata(adata)

# Compute PCA and UMAP embeddings
sc.pp.pca(adata)
sc.pp.neighbors(adata)
sc.tl.umap(adata)
post_embedding_representation = anndata_string(adata)
print_string_diff(
    post_setup_anndata_representation,
    post_embedding_representation,
    diff_title="Embedding diff",
)
print_anndata(adata)

# Create a standard model with default components
model = create_legacy_model1()
print(model)

# Train the model
model.train(
    adata=adata,
    max_epochs=20,  # Use a small number for demonstration
    learning_rate=0.01,
    use_gpu=False  # Set to True if GPU is available
)

# Generate posterior samples
posterior_samples = model.generate_posterior_samples(
    adata=adata,
    num_samples=30  # Use a small number for demonstration
)

# Store results in AnnData
adata = model.store_results_in_anndata(
    adata=adata,
    posterior_samples=posterior_samples
)

# Compute velocity and store it in AnnData
velocity = model.get_velocity(
    adata=adata,
    num_samples=30,
    basis='umap'  # This will be used for embedding
)

# Print available keys in posterior samples
print("Available keys in posterior samples:")
for key in posterior_samples.keys():
    print(f"- {key}: {posterior_samples[key].shape}")
print(model)
```

@@

# Visualize velocity
scv.pl.velocity_embedding_grid(
    adata,
    basis='umap',
    color='gene_0',
    vkey='velocity_pyro'  # Use the velocity key added by get_velocity
    color="gray",
    s=3.5,
    show=False,
    alpha=0.25,
    min_mass=3.5,
    scale=0.35,
    frameon=False,
    density=0.4,
    arrow_size=3,
    linewidth=1,
)
```
